Who mistrust basic science?

### 2024-03-25 paper.Rmd
update draft study 1

-   **RQ1: What is the average science knowledge score (1)?**

-   **RQ2: What is the average acceptance of the scientific consensus (2)?**

-   **RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?**

-   **RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?**

@@

To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by @bruderMeasuringIndividualDifferences2013 and the Single Item Conspiracy Beliefs Scale (SICBS) by @lantianMeasuringBeliefConspiracy2016 (see Appendix \@ref:exp1).

@@

In appendix XX we show that these results hold for our alternative measures of trust and conspiracy thinking. We also include more descriptive statistics, such as knowledge and acceptance by science questions.

Are trust in science and conspiracy thinking, respectively, associated with being more easily convinced of the scientific consensus? In our main analyses, we looked at correlations of acceptance across all observations. One possibility is that the associations between trust in science/conspiracy thinking and acceptance of scientific consensus are explained by science knowledge: People who give the right answer in the first place are more ready to accept the consensus, and trust in science/conspiracy thinking are mostly associated with this knowledge, but not with willingness to accept the consensus. To addressed this potential confound, in a non-preregistered analysis, we restricted our sample to cases where participants gave the wrong answer to the knowledge question. We than calculated the correlation between trust in science and average acceptance rate by participant. We find a slightly smaller negative correlation of acceptance with for conspiracy thinking (r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance`), and only a very small correlation for trust in science (r = `r exp1_descriptives$false_answers_cor_trust_acceptance`). These associations are not statistically significant (see Appendix \@ref:exp1).

The correlations of acceptance with

However, trust/CT is related to basic knowledge, so the fact that it is also related to the post feedback answer could just be a function of that.

## Make a combined plot with all results

@@

These results suggest that even when people do not know the answer to science questions, they tend to mostly accept the scientific consensus. Yet, in `r exp1_descriptives$conditional_acceptance$false$No$share` of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes. In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.

However, conditioning our analyses to initial false knowledge responses, i.e. looking at change of opinion towards the scientific consensus, we find that trust in science is onl

(Experiment 1 Lou write up)

If we ask basic science questions, and we give feedback:
How many people accept the correct answer at the end?
Is that a function of TIS?
Is that a function of CT?

However, trust/CT is related to basic knowledge, so the fact that it is also related to the post feedback answer could just be a function of that.

So we look at who changes their minds. We still see an influence of CT (? do stats) (and of TIS?).

Takeaway: 
Relatively high K in this population
After feedback even higher
But not at ceiling
Relation with TIS/CT, both in K, and in integrating feedback (?)

However, for those who don’t accept the feedback, we don’t know if it’s because they don’t trust us, or they don’t trust science.
We can also try slightly more stringent attention checks (the fact that some people go from good to bad suggests some inattention)

>>>>

**RQ1: What is the average science knowledge score (1)?**

**RQ2: What is the average acceptance of the scientific consensus (2)?**

**RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?**

**RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?**

@@

To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by @bruderMeasuringIndividualDifferences2013 and the Single Item Conspiracy Beliefs Scale (SICBS) by @lantianMeasuringBeliefConspiracy2016 (see Appendix \@ref(exp1)).

@@

In Appendix \@ref(exp1) we show that these results hold for our alternative measures of trust and conspiracy thinking. We also include more descriptive statistics, such as knowledge and acceptance by science questions.

Are trust in science and conspiracy thinking, respectively, associated with being more easily convinced of the scientific consensus? In our main analyses, we looked at correlations of acceptance across all observations. One possibility is that the associations between trust in science/conspiracy thinking and acceptance of scientific consensus are explained by science knowledge: People who give the right answer in the first place are more ready to accept the consensus, and trust in science/conspiracy thinking are mostly associated with this knowledge, but not with willingness to accept the consensus. To addressed this potential confound, in a non-preregistered analysis, we restricted our sample to cases where participants gave the wrong answer to the knowledge question. We than calculated the correlation between trust in science and average acceptance rate by participant. We find a slightly smaller negative correlation of acceptance with for conspiracy thinking (r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance`), and only a very small correlation for trust in science (r = `r exp1_descriptives$false_answers_cor_trust_acceptance`). These associations are not statistically significant (see Appendix \@ref(exp1)).

@@

These results suggest that even when people do not know the answer to science questions, they tend to mostly accept the scientific consensus. Yet, in `r exp1_descriptives$conditional_acceptance$false$No$share` of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes. In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less. However, conditioning our analyses to initial false knowledge responses, i.e. looking at change of opinion towards the scientific consensus, we find that trust in science is onl

### 2024-03-25 paper.tex
update draft study 1

\begin{itemize}
\item
  \textbf{RQ1: What is the average science knowledge score (1)?}
\item
  \textbf{RQ2: What is the average acceptance of the scientific consensus (2)?}
\item
  \textbf{RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?}
\item
  \textbf{RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?}
\end{itemize}

@@

To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by Bruder, Haffke, Neave, Nouripanah, and Imhoff (2013) and the Single Item Conspiracy Beliefs Scale (SICBS) by Lantian, Muller, Nurra, and Douglas (2016) (see Appendix @ref:exp1).

@@

In appendix XX we show that these results hold for our alternative measures of trust and conspiracy thinking. We also include more descriptive statistics, such as knowledge and acceptance by science questions.

Are trust in science and conspiracy thinking, respectively, associated with being more easily convinced of the scientific consensus? In our main analyses, we looked at correlations of acceptance across all observations. One possibility is that the associations between trust in science/conspiracy thinking and acceptance of scientific consensus are explained by science knowledge: People who give the right answer in the first place are more ready to accept the consensus, and trust in science/conspiracy thinking are mostly associated with this knowledge, but not with willingness to accept the consensus. To addressed this potential confound, in a non-preregistered analysis, we restricted our sample to cases where participants gave the wrong answer to the knowledge question. We than calculated the correlation between trust in science and average acceptance rate by participant. We find a slightly smaller negative correlation of acceptance with for conspiracy thinking (r = -0.14), and only a very small correlation for trust in science (r = 0.06). These associations are not statistically significant (see Appendix @ref:exp1).

The correlations of acceptance with

However, trust/CT is related to basic knowledge, so the fact that it is also related to the post feedback answer could just be a function of that.

\hypertarget{make-a-combined-plot-with-all-results}{%
\subsection{Make a combined plot with all results}\label{make-a-combined-plot-with-all-results}}

@@

These results suggest that even when people do not know the answer to science questions, they tend to mostly accept the scientific consensus. Yet, in 23.7 \% of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes. In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.

However, conditioning our analyses to initial false knowledge responses, i.e.~looking at change of opinion towards the scientific consensus, we find that trust in science is onl

(Experiment 1 Lou write up)

If we ask basic science questions, and we give feedback:
How many people accept the correct answer at the end?
Is that a function of TIS?
Is that a function of CT?

However, trust/CT is related to basic knowledge, so the fact that it is also related to the post feedback answer could just be a function of that.

So we look at who changes their minds. We still see an influence of CT (? do stats) (and of TIS?).

Takeaway:
Relatively high K in this population
After feedback even higher
But not at ceiling
Relation with TIS/CT, both in K, and in integrating feedback (?)

However, for those who don't accept the feedback, we don't know if it's because they don't trust us, or they don't trust science.
We can also try slightly more stringent attention checks (the fact that some people go from good to bad suggests some inattention)

>>>>

\textbf{RQ1: What is the average science knowledge score (1)?}

\textbf{RQ2: What is the average acceptance of the scientific consensus (2)?}

\textbf{RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?}

\textbf{RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?}

@@

To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by Bruder, Haffke, Neave, Nouripanah, and Imhoff (2013) and the Single Item Conspiracy Beliefs Scale (SICBS) by Lantian, Muller, Nurra, and Douglas (2016) (see Appendix \ref{exp1}).

@@

In Appendix \ref{exp1} we show that these results hold for our alternative measures of trust and conspiracy thinking. We also include more descriptive statistics, such as knowledge and acceptance by science questions.

Are trust in science and conspiracy thinking, respectively, associated with being more easily convinced of the scientific consensus? In our main analyses, we looked at correlations of acceptance across all observations. One possibility is that the associations between trust in science/conspiracy thinking and acceptance of scientific consensus are explained by science knowledge: People who give the right answer in the first place are more ready to accept the consensus, and trust in science/conspiracy thinking are mostly associated with this knowledge, but not with willingness to accept the consensus. To addressed this potential confound, in a non-preregistered analysis, we restricted our sample to cases where participants gave the wrong answer to the knowledge question. We than calculated the correlation between trust in science and average acceptance rate by participant. We find a slightly smaller negative correlation of acceptance with for conspiracy thinking (r = -0.14), and only a very small correlation for trust in science (r = 0.06). These associations are not statistically significant (see Appendix \ref{exp1}).

@@

These results suggest that even when people do not know the answer to science questions, they tend to mostly accept the scientific consensus. Yet, in 23.7 \% of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes. In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less. However, conditioning our analyses to initial false knowledge responses, i.e.~looking at change of opinion towards the scientific consensus, we find that trust in science is onl
